Explorative data structure comparison with application to \ldots

### 2017-03-02 article/article_v0.01.tex
Emphasized difference between correlation and "covariance of standardized variables".

In this paper we propose three diagnostic plots for comparing the loading patterns in two different datasets. These three plots will be referred to as the \emph{CumEigenPlot}, the \emph{HairPlot} and the \emph{PancakePlot}. In order to describe these plots we consider two different datasets with $n_1$ and $n_2$ observations, respectively, but in the same $d$ variables. We assume that the variables have been standardized separately within the two datasets to have zero mean and unit standard deviation \hl{(justification for this choice should be further elaborated. And also: Something about what to do if not all variables are numerical.)}. The empirical covariance matrices for the two datasets are called $S_1$ and $S_2$, respectively. Due to the standardization these are the same as the empirical correlation matrices. Finally, let $\lambda_{i1} \ge \dotsm \ge \lambda_{id}$ and $\eta_{i1},\dotsc,\eta_{id}$ be the variation components and the loadings for dataset $i=1,2$.

>>>>

In this paper we propose three diagnostic plots, referred to as the \emph{CumEigenPlot}, \emph{HairPlot} and \emph{PancakePlot}, for comparing the loading patterns in two datasets. In order to describe these plots we consider two different datasets in the same $d$ variables and with $n_1$ and $n_2$ observations, respectively. For dataset $i=1,2$ let $S_i \in \RR^{d \times d}$ be the empirical correlation matrix, and let $\lambda_{i1} \ge \dotsm \ge \lambda_{id} \ge 0 $ and $\eta_{i1},\dotsc,\eta_{id} \in \RR^{d}$ be the corresponding variation components and loadings. The correlation matrices correspond to the covariance matrices for the variables after these have been standardized to unit standard deviation separately within the two datasets.

\hl{To do: justification for using correlation matrices should be further elaborated. And also: Something about what to do if not all variables are numerical. Using correlation matrices is not the same as assuming standadized variables; there is a difference in the permutations.}
